Advancing spatially explicit fisheries management with age-structured species distribution models

### 2025-01-28 paper.qmd
reference and text

To illustrate potential benefits to fisheries management, we used output from our models to generate spatiotemporal predictions of age-1 sablefish (from age-0 sablefish the year prior). \[Need some text here on background why sablefish as bycatch is important. In 2017 for example, the at-sea whiting fishery had 50 mt of sablefish quota set aside in the area north of 36 deg N, but the fleet caught more than 3x as much\]. Gridded predictions from the Poisson GLM represent the expected number of fish per unit of effort in each location. To turn these into a potentially more useful measure of risk, we calculate the predicted total age-1 sablefish catch per unit of effort (CPUE) within a given radius of major fishing ports on the US west coast. For consistency with previous work \[TODO: cite ESR\] we used a radius of 232km. These predictions of port-level CPUE were then summarized as annual time series (1-step ahead forecasts).

>>>>

To illustrate potential benefits to fisheries management, we used output from our models to generate spatiotemporal predictions of age-1 sablefish (from age-0 sablefish the year prior). Sablefish are part of the DTS complex, which includes Dover sole *Microstomus pacificus*, shortspine *Sebastomus alascanus,* longspine thornyheads *S. altivelis*, and sablefish. This complex forms and inportant fishery on the West Cast, and by-catch of sablefish can impact quotas for the other three species, as well as the at-sea Hake fishery. For exmaple, in 2017, the at-sea hake fishery had 50 mt of sablefish quota set aside in the area north of 36 °N, but the fleet caught more than 3x as much, promting in-season warnings to alter fishing behavior.

Gridded predictions from the Poisson GLM represent the expected number of fish per unit of effort in each location. To turn these into a potentially more useful measure of risk, we calculate the predicted total age-1 sablefish catch per unit of effort (CPUE) within a given radius of major fishing ports on the US west coast. For consistency with previous work [@leising2024cciea], we used a radius of **282km** **\[was listed as 232 (my fault) but should be 282 to be consistent with the ESR\]**. These predictions of port-level CPUE were then summarized as annual time series (1-step ahead forecasts).
